Synthesis of Linear Functional Programs

### 2021-05-09 article/article.tex

@@

We've already implemented a parser and a type-checker for a linear
$\lambda$-calculus using the resource management technique described
in previous section -- more than
good preparation for the development of the synthesizer, they will be
useful to \emph{validate} the synthesized functions, and to enable
the synthesis of partial programs.
%
The implementations have been carried out in Haskell since it's easier
to model and implement inference rules in a functional setting.

\mypara{Evaluation} Validation and evaluation of our work is fairly
simple. Simple questions such as \emph{does it typecheck}? \emph{can
  we synthesize}? \emph{how fast}?  are the core of our evaluation. In
more concrete terms, we can compare our times to state of the art
synthetizers, present a list of working examples of programs synthesized
by our framework, and corresponding benchmarks.

@@

% \section{Bonus}

% \[
%   \begin{array}{lcl}
%     A, B & ::= & A \tensor B \mid A \lolli B \mid A \with B \mid A
%                  \oplus B \mid \bang A \mid \one \mid \dots\\[1ex]
%     M,N & ::= & \lambda x{:}A.M\\
%          & \mid & M\,N\\
%          & \mid & (M \tensor N)\\
%          & \mid & \llet{x\tensor y = M}{N} \\
%          & \mid & x \\
%          & \mid & \dots\\
%     \end{array}
% \]

% \[
%   \infer[$(\lolli\! I)$]
%   {\Delta , x{:}A \vdash M : B}
%   {\Delta \vdash \lambda x {:} A . M : A \lolli B}
%   \quad
%   \infer[$(\lolli\! E)$]
%   {\Delta_1 \vdash M : A \lolli B \and \Delta_2 \vdash N : A}
%   {\Delta_1, \Delta_2 \vdash M\,N : B}
% \]

% \[
%   \infer*[left=($\tensor I$)]
%   {\Delta_1 \vdash M : A \and \Delta_2 \vdash N : B}
%   {\Delta_1 , \Delta_2 \vdash (M \tensor N) : A \tensor B}
%   \quad
%   \infer*[right=($\tensor E$)]
%   {\Delta_1 \vdash M : A \tensor B \and \Delta_2 , x{:}A, y{:}B\vdash
%     N : C }
%   {\Delta_1 , \Delta_2\vdash \llet{x\tensor y = M}{N} : C }
% \]

>>>>

This is the theory we're interested in and have studied so far,
which will allow us to build the actual synthetizer. From here,
our work plan gives an overview of the steps we'll take.

@@

We've already implemented a parser and a type-checker for a linear
$\lambda$-calculus using the resource management technique described
in the previous section -- more than
good preparation for the development of the synthesizer, they will be
useful to \emph{validate} the synthesized functions, and to enable
the synthesis of partial programs.
%
The implementations have been carried out in Haskell since it's easier
to model and implement inference rules in a functional setting.

\mypara{Step by step} Our work so far has been constantly accompanied
by somewhat dense theoretical material, state of the art,
which we must understand to overcome challenges in these fields,
and build from the start a more scalable synthetizer -- we want
to display concrete results to prove the feasability of synthetising
linear types, and thus possibility for future application to related fields,
such as session types for message passing concurrency.

Apart from the theory, when we started studying type systems, we created
a simple type-checker for the simply-typed $\lambda$-calculus.
While learning about linear type systems, and the simply-typed linear
$\lambda$-calculus, we developed a parser, a desugar module, and an
evaluation module to write simple programs in the language;
culminating in writing the type-checker for linear types. At this point,
where we stand as of now, we can parse, type-check, and evaluate this
small linear functional language.
The next step is to start developing the actual synthetizer, embedding it
with all the techniques we've seen. Following work will be done to
augment usability, expand the language domain, and add features such as the
ones mentioned above.

\mypara{Evaluation} Validation and evaluation of our work is fairly
easy. Simple questions such as \emph{does it typecheck}? \emph{can
  we synthesize}? \emph{how fast}?  are the core of our evaluation. In
more concrete terms, we can compare our times to state of the art
synthetizers, present a list of working examples of programs synthesized
by our framework, and corresponding benchmarks.

@@
